---
title: "Visualize and plot interactive Bezier Curves with Beziervizer"
author: Ziyan Xiong
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualize and plot interactive Bezier Curves with Beziervizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

Beziervizer is an R package for plotting interactive [Bézier curves](https://en.wikipedia.org/wiki/B%C3%A9zier_curve). Users can move the control points of plotted result at will and the curve will transform accordingly. This package also includes functions to demonstrate the properties of Bézier curves such as De Casteljau's algorithm and rational Bézier curves in an interactive way.

# Plot an interactive bezier curve

The fastset way to get an interactive Bézier curve is calling function `plot_cubic_bezier()` as follows:

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.center {
  display: table;
  margin-right: auto;
  margin-left: auto;
}
```

```{r setup}
library(Beziervizer)
```


```{r cubic_bezier, results='hide', message=FALSE}
#Plot a cubic bezier curve with default control points
plot_cubic_bezier()
```

<div class = 'center'>
```{r cubic_bezier, out.width="400px", out.height="400px", echo = F}
```
</div>

In the interactive plot above, users can drag the control points (the blue points) at will and observe how the black Bézier curves transform according to these control points.

Users can also use their own control points to initialize the Bézier curves. The `plot_cubic_bezier()` also accepts dataframe and matrix as input. Here is a dataframe input example below.

```{r user_input, results='hide', message=FALSE}
#Plot a bezier curve with a dataframe as control points
x = c(200, 400, 100, 300)
y = c(500, 300, 300, 500)
control_points = data.frame(x,y)
plot_cubic_bezier(control_points)
```

<div class = 'center'>
```{r user_input, out.width="400px", out.height="400px", echo = F}
```
</div>

One thing about the coordinates of control points is that the shape of a Bézier curve formed by a set of control points is determinded by the relative position of control points. Therefore, the value of control points' coordinates is less important. The function `plot_cubic_bezier()` will resize the input control points to suitable size to display automatically. The coordinates in the example below will generate the same Bézier curve as the example above.

```{r , results='hide', message=FALSE}
#Plot a bezier curve with a dataframe as control points
x = c(0, 2, -1, 1)
y = c(5, 3, 3, 5)
control_points = data.frame(x,y)
plot_cubic_bezier(control_points)
```

## De Casteljau's Algorithm

[De Casteljau's Algorithm](https://en.wikipedia.org/wiki/De_Casteljau%27s_algorithm) introduces a way to draw Bézier curves without calculating the coordinates of points on the Bézier curves explicitly. The slider in the bottom of the plot controls the parameter `t` in calculating the position of points on a Bézier curve.

```{r de_castlejau, results='hide', message=FALSE}
#Plot a bezier curve with  with a matrix as control points
x = c(100, 200, 300, 400)
y = c(500, 200, 300, 600)
control_points = matrix(c(x,y),length(x),2)
plot_cubic_bezier(control_points)
```

<div class = 'center'>
```{r de_castlejau, out.width="400px", out.height="400px", echo = F}
```
</div>

Let's denote the blue control points as $P_{00}, P_{10}, P_{20}, P_{30}$, the yellow points as $P_{01}, P_{11}, P_{21}$, the red points as $P_{02}, P_{12}$ and finally the black point on Bézier curve as $P_{03}$. De Casteljau's Algorithm get $P_{03}$ from $P_{00}, P_{10}, P_{20}, P_{30}$ by performing following operation recursively.

\[ P_{ij} = P_{i(j-1)}(1-t) + P_{(i+1)(j-1)}t\]

The animation of this process in shown in the interactive plot above. Users can drag the handle to see the how the black point on the Bézier curve will move following the change of `t`. Users can also click the Play button to see it increases automatically.

An important property of the intermediate points generated in the process of De Casteljau's Algorithm is that $P_{00},P_{01},P_{02},P_{03}$ form the control points for the Bezier curve from $P_{00}$ to $P_{03}$ and $P_{03},P_{10},P_{21},P_{30}$ form the control points for the Bezier curve from $P_{03}$ to $P_{30}$. To make it easier to see, if you find the first and last points of each color and connect them following the order of "blue, yellow, red, black", you will get the new control points for the front and back part of the Bezier curve divided by the black point respectively. This property is the key to calculate high order Bézier curve effectively in next part.

# Plot a high order interactive bezier curve
